Characteristic polynomial of p-adic matrices

### 2017-01-10 charpoly/charpoly.tex
Minor improvements: shear matrices and rec. poly.

\begin{prop} 
Algorithm 1 computes $H$ and $P$ realizing an approximate Hessenberg form of $M.$
$P$ is exact over finite extensions of $\mathbb{Q}_p$ and $k((X))$, and the computation is in $O(n^3)$ operations in $K$ at precision the maximum precision of a coeffient in $M.$
\end{prop}
\begin{proof}
Let us assume that $K$ is a finite extensions of $\mathbb{Q}_p$ or $k((X)).$
Inside the nested \textbf{for} loop, if we want to eliminate $p^{u_y} \varepsilon_y+O(p^{n_y})$ with pivot $p^{u_x} \varepsilon_x+O(p^{n_x}),$
with the $\varepsilon$'s being units,
the coefficient of the transvection matrix is the lift (in $\mathbb{Z}, $  $k[X]$ or adequate extension) of $p^{u_y-u_x} \varepsilon_y \varepsilon_x^{-1} \mod \pi^{\min (n_x,n_y)}.$
Exactness follows directly. Over other fields, we can not lift, but the computations are still valid.
The rest is clear.
\end{proof}

@@

The second ingredient we need is the reciprocal polynomial.
We extend its definition to matrices of polynomials.
\begin{deftn}
Let $A \in M_n(K[X]).$ We denote by $A^{rec}$ the matrix
whose coefficients are the reciprocal polynomials of the
coefficients of $A.$ 
\end{deftn}
We then have the following result :
\begin{lem}
Let $M \in M_n(K).$ Then:
\begin{eqnarray*}
com(1-XM)^{rec}=&com(X-M), \\
(\chi_M \times 1)^{rec}=&(1-XM) com(1-XM).\\
\end{eqnarray*}
\end{lem}

@@

\begin{theo}
Let $H \in M_n(O_K)$ be an approximate Hessenberg matrix.
Then, using Algorithm 2, one can compute $\com (X -H)$ in 
$\softO (n^3)$ operations in $O_K$ at the precision given by $H.$
\end{theo}
\begin{proof}
Firstly, the operations of the lines 2 and 3 are in $\softO (n^3)$ operations
in $O_K$ at the precision given by $H.$
Indeed, since $H$ is an approximate matrix, when we use $U_{i,i}$ as pivot:
on its columns, only $U_{i+1,i}$ can be a non zero-approximate coefficient.
As a consequence, when performing this column-pivoting, only two rows ($i$ and
$i+1$) lead to operations in $O_K [[K]]$ other than checking precision.
Hence, line 3 costs $\softO (n^2)$ for the computation of $U.$
Following line 1, the computation of $Q^{-1}$ is done by operations on rows, starting from the identity matrix.
Coefficient for the transvection is the opposite of the one defined by pivoting
and the indices for the receiving and giving rows are swapped compared to the operations on columns.
Thus, we can show that it requires no additional operation in $O_K[[X]].$ $Q^{-1}$ is just filled as
an upper-triangular matrix.
This proves that the resulting cost for lines 2 and 3 is indeed in $\softO (n^3)$ operations.

@@

Thanks to the fact that the $P$ only corresponds to the product of $n-1$ transvections, the
product on line 8 is in $\softO (n^3).$
We emphasize that no division has been done throughout the algorithm.
Line 9 is costless, and the result is then proved.
\end{proof}

>>>>

\begin{prop} 
Algorithm 1 computes $H$ and $P$ realizing an approximate Hessenberg form of $M.$
$P$ is exact over finite extensions of $\mathbb{Q}_p$ and $k((X))$, and the computation is in $O(n^3)$ operations in $K$ at precision the maximum precision of a coeffient in $M.$
\end{prop}
\begin{proof}
Let us assume that $K$ is a finite extensions of $\mathbb{Q}_p$ or $k((X)).$
Inside the nested \textbf{for} loop, if we want to eliminate $p^{u_y} \varepsilon_y+O(p^{n_y})$ with pivot $p^{u_x} \varepsilon_x+O(p^{n_x}),$
with the $\varepsilon$'s being units,
the corresponding coefficient of the corresponding shear matrix is the lift (in $\mathbb{Z}, $  $k[X]$ or adequate extension) of $p^{u_y-u_x} \varepsilon_y \varepsilon_x^{-1} \mod \pi^{\min (n_x,n_y)}.$
Exactness follows directly. Over other fields, we can not lift, but the computations are still valid.
The rest is clear.
\end{proof}

@@

The second ingredient we need is the reciprocal polynomial.
We extend its definition to matrices of polynomials.
\begin{deftn}
Let $P \in K[X]$ be of degree $d \in \mathbb{N}_{>0}.$ 
We define its reciprocal polynomial as $P^{rec}=X^d P \left( 1/X \right).$
Let $A \in M_n(K[X]).$ We denote by $A^{rec}$ the matrix
whose coefficients are the reciprocal polynomials of the
coefficients of $A.$ 
\end{deftn}
We then have the following result :
\begin{lem}
Let $M \in M_n(K).$ Then:
\begin{eqnarray*}
com(1-XM)^{rec}=&com(X-M), \\
(\chi_M \times 1)^{rec}=&(1-XM) com(1-XM).\\
\end{eqnarray*}
\end{lem}

@@

\begin{theo}
Let $H \in M_n(O_K)$ be an approximate Hessenberg matrix.
Then, using Algorithm 2, one can compute $\com (X -H)$ in 
$\softO (n^3)$ operations in $O_K$ at the precision given by $H.$
\end{theo}
\begin{proof}
Firstly, the operations of the lines 2 and 3 are in $\softO (n^3)$ operations
in $O_K$ at the precision given by $H.$
Indeed, since $H$ is an approximate matrix, when we use $U_{i,i}$ as pivot:
on its columns, only $U_{i+1,i}$ can be a non zero-approximate coefficient.
As a consequence, when performing this column-pivoting, only two rows ($i$ and
$i+1$) lead to operations in $O_K [[K]]$ other than checking precision.
Hence, line 3 costs $\softO (n^2)$ for the computation of $U.$
Following line 1, the computation of $Q^{-1}$ is done by operations on rows, starting from the identity matrix.
Corresponding coefficient for the corresponding shear matrix
 is the opposite of the one defined by pivoting
and the indices for the receiving and giving rows are swapped compared to the operations on columns.
Thus, we can show that it requires no additional operation in $O_K[[X]].$ $Q^{-1}$ is just filled as
an upper-triangular matrix.
This proves that the resulting cost for lines 2 and 3 is indeed in $\softO (n^3)$ operations.

@@

Thanks to the fact that the $P$ only corresponds to the product of $n-1$ 
shear matrices, the
product on line 8 is in $\softO (n^3).$
We emphasize that no division has been done throughout the algorithm.
Line 9 is costless, and the result is then proved.
\end{proof}
